\TheTitle

### 2016-07-25 linvis.tex
Small eidtorial changes in many places.
- Removed empty space in some of the images
- Edited text in few locations
- updated makefile to make diffs

\newcommand{\TheTitle}{Merge-tree: Visualizing the integration of commits into Linux}
\newcommand{\TheAuthors}{Evan Wilde, Daniel German}
\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
\newcommand{\TheSubject}{Digesting large amounts of commit data}
\newcommand{\TheKeywords}{Linux, git, data structures, tree data structures}

@@

Git uses a directed acyclic graph (DAG) as its main data model. In this model, a
commit has one or more ordered parents (ancestors), except the root commits of a
repository which do not have ancestors--Linux git repository has two such commits.
Commits are divided into merge commits, merging two or more branches, and non-merge
commits. Non-merge commits have only one parent, while merge commits have two or
more. The order of the parents matter: the first parent is the branch in which the
merge is being done, while the rest indicate the branches being merged (Git allows
merging multiple branches simultaneously). The second parent is the first branch
merged, the third parent is the second branched merged, etc.

Once a commit is created, it is never changed. Git allows operations to alter
commits or reorder them, but it changes the commit ID in the process, effectively
making a new commit. This makes commits unable to record the traversal of merges
from the commit to the merge into the master branch.

@@

\begin{figure}[htbp]
        \centering
        {\tiny
        \begin{verbatim}
Merge: 8cbd84f fd8aa2c
Author: Linus Torvalds <torvalds@linux-foundation.org>
Date:   Tue Aug 10 15:38:19 2010 -0700

@@

* 'for-linus' of git://neil.brown.name/md: (24 commits)
md: clean up do_md_stop
[... edited for the sake of space]
md: split out md_rdev_init
md: be more careful setting MD_CHANGE_CLEAN
md/raid5: ensure we create a unique name for kmem_cache when mddev has no gendisk
...
        \end{verbatim}}
        \caption{Example of how merges record a subset of commits being merged. The
                commit only shows the first 20 one-line log messages for the 24
                non-merge commits it merged. The ending ``...'' is part of the log
                and represents that other commits were merged.}
        \label{fig:sampleMerge}
\end{figure}

@@

To navigate and inspect the merge-tree view of the kernel we created a web-based
tool called \tool. Creating a web-based tool enables users to use the system without
having to install additional software or store a large database, making it more
accessible, more easily maintainable, and platform independent. \tool uses the
following mechanisms to reach our goals of better navigation and better explanation
of the changes.

@@

Searching (depicted in Figure \ref{fig:search}) allows a user to filter commits and
merges that are irrelevant. The search mechanism breaks down the results by release
version. A user can further narrow down the search by specifying a range of dates in
which such commits were merged by Linus into the master branch, not when the commits
were created. This distinction is important. We have observed commits that have
taken years to arrive into the master repository after they were originally
committed.

@@

If the user is searching by commit ID, ID must be specified from the first character
to the last character. For example, the commit
`c267548755a184ef97301071300c1739a564e135' is returned when the user searches for a
commit ID of `c26' in the 3.10 Linux kernel. This commit is a non-merge commit
performed in one of the largest merge-trees of the kernel; however, this feature
will work with any commit ID.

@@

In the search results, the user is presented with the one-line log message preview,
the author's name and email, the date the commit was authored, and the date the
commit was last committed. The commit date and author date will usually be the same;
however, if the commit has been rebased or cherry-picked into a merge that
eventually is merged into the kernel, the commit date and authored date will be
different. We can see this in the first entry in Figure~\ref{fig:results}.

@@

The modules tab (Figure~\ref{fig:modules}) shows the modules that are contained
within the commit. Modules are not natively recognized by Git, and are not going to
be present in all repositories. In the Linux repository, authors put the module they
are working on in the one-line preview of the log-message; for example: the log
\emph{gcov: add support for FCC 4.9} has updated the \emph{gcov} module (the
coverage testing tool of the kernel). We heuristically extract the module by taking
all text in the log preview of commits until the first colon.  Modules are logical
partitions of the information in the kernel. Depending on where the author was
working, modules can be general, such as ``bluetooth'' and ``wireless'', or can be
quite specific for individual hardware, such as ``ath9k\_hw'' and ``wl1251''. In a
few cases, the author of a commit does not correctly follow this format and the
heuristic approach fails. As with the Files panel, non-merge commits show their
corresponding information, but for merge commits we aggregate all the modules
changed in all the commits that are part of that merge.  The output of this view is
shown in Figure~\ref{fig:modules}.

@@

The tree can be further extended to include the collapsing of merge nodes, only
showing parts of the tree that are of interest to the user. Furthermore, the user can
zoom the tree by scrolling the mouse wheel to see more of the topology of the tree,
or to see the details of a given part of the tree.

@@

\begin{tabular}{ccl}
        Field name      & Data type & Description\\\hline
        \verb|cid|      & string    & Commit ID\\
        \verb|name|     & string    & One-line log preview\\
        \verb|mlinus|   & string    & Root merge commit ID\\
        \verb|author|   & string    & Author name and email\\
        \verb|mnext|    & string    & Parent merge commit ID\\
        \verb|children| & object    & Object of tree objects\\
\end{tabular}

@@

From a developers point of view, the are two major disadvantages of the DAG model of
Git: a) its edges point backwards, i.e. commits point to their ancestor, not to the
commit that succeeds them on the path to being integrated; and b) the DAG contains
many more edges that are necessary to understand how integration occurred. We
addressed these two issues with the creation of the merge-tree model of the DAG of
Linux. Effectively the merge-tree view recovers the details of each merged branch
into the master branch.

In our experience, no other Git repository reaches the level of DAG complexity that
Linux has. In most Git repositories, merges are not nested, all merges merge
directly into the master repository, and these merges only merge few commits. Even
in these simplified cases, the merge-tree can provide a valuable summary of how
commits are integrated into the master branch, specifically since the time of
integration may be very different from the time the commits were authored.

@@

The biggest challenge is computing the merge-tree of the Linux kernel repository. It
is likely that our heuristic for computing the merge-tree will not work with other
repositories. The heuristic requires that there are no fast-forward merges into the
master branch of the repository (a majority of repositories), and that there are no
foxtrot merges. We are able to use our heuristic with the Linux kernel because of
the strict integration model imposed by Linus. To validate the results of the
heuristic, we use the continuous mining technique described in \cite{German2015}.
The continuous mining of the repository requiring foresight and planning.

@@

This merge is relatively small, containing only four sub-merges, three of which
only contain a single commit, and the last containing two commits. With our
tool, we are able to immediately see what section of the kernel these commits and merges pertain to. The DAG view of this provides almost no explanation,
furthermore, users must work to determine where this merge ends and the next one
begins.

@@

Our tool shows promising results in working toward our goals, building a tool that
can more easily navigate the repository, and provide clearer explanations of the
changes made. The filtering by searches and the various tree visualizations enable
users to easily navigate through the kernel, finding the commits and merges that
pertain to what they are interested in. The tree visualizations are able to both
serve as a navigation piece, providing simpler navigation through the kernel, but
also demonstrate which commits contribute to which merges, providing further
explanation over the DAG. The tree model further improves the explanation by enabling
the tool to aggregate metadata at the merges instead of requiring a user to manually
aggregate the information. Our tool currently aggregates file information and the
modules edited, but can be extended to include authorship information and aggregated
patch information.

>>>>

\newcommand{\TheTitle}{Merge-tree: Visualizing the Integration of Commits into Linux}
\newcommand{\TheAuthors}{Evan Wilde, Daniel German}
\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
\newcommand{\TheSubject}{Digesting large amounts of commit data}
\newcommand{\TheKeywords}{Linux, git, data structures, tree data structures}

@@

Git uses a directed acyclic graph (DAG) as its main data model. In this model, a
commit has one or more ordered parents (ancestors), except the root commits of a
repository which do not have ancestors--Linux's git repository has two such commits.
Commits are divided into merge commits, merging two or more branches, and non-merge
commits. Non-merge commits have only one parent, while merge commits have two or
more. The order of the parents matter: the first parent is the branch in which the
merge is being done, while the rest indicate the branches being merged (Git allows
merging multiple branches simultaneously). The second parent is the first branch
merged, the third parent is the second branched merged, etc.

Once a commit is created, it is never changed. Git allows operations to alter
commits or reorder them, but it changes the commit ID in the process, effectively
replacing it with a new commit. This makes commits unable to record the traversal of merges
from the commit to the merge into the master branch.

@@

\begin{figure}[htbp]
        \centering
        {\fontsize{7}{9}
        \begin{verbatim}
Merge: 8cbd84f fd8aa2c
Author: Linus Torvalds <torvalds@linux-foundation.org>
Date:   Tue Aug 10 15:38:19 2010 -0700

@@

* 'for-linus' of git://neil.brown.name/md: (24 commits)
md: clean up do_md_stop
[... edited for the sake of space]
md: split out md_rdev_init
md: be more careful setting MD_CHANGE_CLEAN
md/raid5: ensure we create a unique name for kmem_cache...
...
        \end{verbatim}}
        \caption{Example of how merges record a subset of commits being merged. The
                commit only shows the first 20 one-line log messages for the 24
                non-merge commits it merged. The ending ``...'' is part of the log
                and represents that other commits were merged.}
        \label{fig:sampleMerge}
\end{figure}

\textbf{rewrite}

@@

To navigate and inspect the merge-tree view of the kernel we created a web-based
tool called \tool. Creating a web-based tool enables users to use the system without
having to install additional software or store a large database, making it more
accessible, more easily maintainable, and platform independent. \tool uses the
following mechanisms to reach our goals of better navigation and better explanation
of the selected changes.

@@

Searching (depicted in Figure \ref{fig:search}) allows a user to filter commits and
merges that are irrelevant. The search mechanism breaks down the results by release
version. A user can further narrow down the search by specifying a range of dates in
which such commits were merged by Linus into the master branch---not when the commits
were created (author date and/or commit date). This distinction is important. We have observed commits that have
taken years to arrive into the master repository after they were originally
created.

@@

If the user is searching by commit ID, the ID can be specified by using any of its unique prefixes. For example, the commit
\mycode{c267548755a184ef97301071300c1739a564e135} is returned when the user searches for a
commit ID of \mycode{c26} in the 3.10 Linux kernel. % This commit is a non-merge commit
% performed in one of the largest merge-trees of the kernel; however, this feature
% will work with any commit ID.

@@

In the search results (seen in Figure~\ref{fig:results}), the user is presented with the one-line log message preview,
the author's name and email, the date the commit was authored, and the date the commit was last committed.

@@

The modules tab (Figure~\ref{fig:modules}) shows the modules that are contained
within the commit. Modules are not natively recognized by Git, and are not going to
be present in all repositories. In the Linux repository, authors put the module they
are working on in the one-line summary of the log-message; for example: the log
\emph{gcov: add support for FCC 4.9} has updated the \emph{gcov} module (the
coverage testing tool of the kernel). We heuristically extract the module by taking
all text in the log summary of commits until the first colon.  Modules are logical
partitions of the information in the kernel. Depending on where the author was
working, modules can be general, such as ``bluetooth'' and ``wireless'', or can be
quite specific for individual hardware, such as ``ath9k\_hw'' and ``wl1251''. In a
few cases, the author of a commit does not correctly follow this format and the
heuristic approach fails. As with the Files panel, non-merge commits show their
corresponding information, but for merge commits we aggregate all the modules
changed in all the commits that are part of that merge.  The output of this view is
shown in Figure~\ref{fig:modules}.

@@

The tree can be displayed with some merges nodes collapsed, hence
showing only parts of the tree that are of interest to the user. Furthermore, the user can
zoom the tree by scrolling the mouse wheel to see more of the topology of the tree,
or to see the details of a given part of the tree.

@@

\begin{tabular}{ccl}
        Field name      & Data type & Description\\\hline
        \verb|cid|      & string    & Commit ID\\
        \verb|name|     & string    & One-line log summary\\
        \verb|mlinus|   & string    & Root merge commit ID\\
        \verb|author|   & string    & Author name and email\\
        \verb|mnext|    & string    & Parent merge commit ID\\
        \verb|children| & object    & Object of tree objects\\
\end{tabular}

@@

From a developers point of view, the are two major disadvantages of the DAG model of
Git: a) its edges point backwards, i.e. commits point to their ancestor, not to the
commit that succeeds them on the path to being integrated; and b) the DAG contains
many more edges that are necessary to understand how integration occurred. We
addressed these two issues with the creation of the merge-tree model of the DAG of
Linux. Effectively the merge-tree view recovers the details of how each branch---and the commits they contain---was merged
into the master branch.

In our experience, no other Git repository reaches the level of DAG complexity that
Linux has. In most Git repositories, most merges are not nested---most merges merge
directly into the master repository, and these merges only merge few commits. Even
in these simplified cases, the merge-tree can provide a valuable summary of how
commits are integrated into the master branch, specifically since the time of
integration may be very different from the time the commits were authored.

@@

The biggest challenge is computing the merge-tree of any given repository. It
is likely that our heuristic for computing the merge-tree will not work with other
repositories. The heuristic requires that there are no fast-forward merges into the
master branch of the repository (a majority of repositories), and that there are no
foxtrot merges (this is practice that starts be considered desirable by git users\footnote{See
  \url{http://devblog.nestoria.com/post/98892582763/maintaining-a-consistent-linear-history-for-git} and
  \url{https://developer.atlassian.com/blog/2016/04/stop-foxtrots-now/}}).
We are able to use our heuristic with the Linux kernel because of
the strict integration model imposed by Linus. To validate the results of the
heuristic, we use the continuous mining technique described in \cite{German2015}.
The continuous mining of the repository requiring foresight and planning.

@@

Even though this merge is relatively small, containing only four sub-merges, three of which
only contain a single commit, and the last containing two commits, it is already hard to visualize in Gitk. 
With \tool we are able to immediately see what section of the kernel these commits and merges pertain to. The DAG view of this provides almost no explanation,
furthermore, users must work to determine where this merge ends and the next one
begins.

@@

Our tool, \tool, shows promising results in working toward our goals, building a tool that
can more easily navigate the repository, and provide clearer explanations of the
changes made. The filtering by searches and the various tree visualizations enable
users to easily navigate through the kernel, finding the commits and merges that
pertain to what they are interested in. The tree visualizations are able to both
serve as a navigation piece, providing simpler navigation through the kernel, but
also demonstrate which commits contribute to which merges, providing further
explanation over the DAG. The tree model further improves the explanation by enabling
the tool to aggregate metadata at the merges instead of requiring a user to manually
aggregate the information. Our tool currently aggregates file information and the
modules edited, but can be extended to include authorship information and aggregated
patch information.
